---
title: "Liquid Viscosity of Hydrocarbon Prediction Model"
output:
  html_notebook: default
  pdf_document: default
---

<p>Introductory text here.</p>

## Setup
<p>Start by loading in the libraries that will be useful.</p>
```{r}
library(tidyverse)
library(car)
library(lmtest)
```

## Model Setup
<p>Now we need to create the data set that we'll us to determine our regression coefficients. The goal is to reproduce the coefficients published by <em>Van Velzen et al</em>. We need to extract the published values for $N_E$ and $T_0$ from the appendices. Then, those data will be used to run a few different polynomial regression models on the data. The regression models will all use a simply monomial basis in a Vandermonde matrix.</p>

```{r}
N_exp = c(0.79, 1.91, 3.13, 4.07, 4.91, 5.91, 7.04, 7.97, 9.05, 9.99, 
       11.04, 11.99, 12.99, 13.98, 15.02, 16.00, 17.01, 18.01, 19.00, 19.98)

T0 = c(57.60, 95.57, 133.41, 160.20, 182.48, 207.09, 232.53, 251.71, 272.12, 288.37, 
       305.01, 318.78, 332.10, 344.21, 355.92, 366.11, 375.90, 385.00, 393.54, 401.67)

B_exp = c(114.14, 156.60, 222.67, 265.84, 313.66, 362.79, 436.73, 473.70, 525.56, 558.61,
      605.50, 631.63, 664.10, 689.85, 718.51, 738.30, 757.88, 777.40, 793.62, 811.29)

df_raw = data.frame(N_exp, T0, B_exp)

N = N_exp
N2 = N^2
N3 = N * N2
N4 = N * N3
```
<p>Perform regressions to determine the coefficients for $T_0$.</p>
```{r}
fit_quadratic_T0 = lm(T0 ~ N_exp + I(N_exp^2), data = df_raw)
summary(fit_quadratic_T0)

fit_cubic_T0 = lm(T0 ~ N_exp + I(N_exp^2) + I(N_exp^3), data = df_raw)
summary(fit_cubic_T0)

fit_quartic_T0 = lm(T0 ~ N_exp + I(N_exp^2) + I(N_exp^3) + I(N_exp^4), data = df_raw)
summary(fit_quartic_T0)
```
### Notes
<p>Testing different polynomial regressions, the cubic fit is the best. The $R^2$ values for all flavours were very high. The quartic fit is redundant, the extra term does not contribute anything to the fit. Give that the $R^2$ value for the cubic model is 1, this is no surprise at all.<p>
<p>The coefficients for the cubic model are:</p>
<p>$\beta_0 = 28.859$, $\beta_1 = 37.439$, $\beta_2 = -1.3547$, $\beta_3 = 0.02076$</p>
<p>These coefficients reproduce those published by <em>Van Velzen et al.</em> in Eq. 7 (pg 7).</p>

```{r}
coef(fit_cubic_T0)
```
<p>Perform regressions to determine the coefficients for $B$.</p>

```{r}
fit_quadratic_B = lm(B_exp ~ N_exp + I(N_exp^2), data = df_raw)
summary(fit_quadratic_B)

fit_cubic_B = lm(B_exp ~ N_exp + I(N_exp^2) + I(N_exp^3), data = df_raw)
summary(fit_cubic_B)

fit_quartic_B = lm(B_exp ~ N_exp + I(N_exp^2) + I(N_exp^3) + I(N_exp^4), data = df_raw)
summary(fit_quartic_B)
```
```{r}
coef(fit_cubic_B)
```

<p>The B regression coefficients do not match those published by <em>Van Velzen et al.</em> but they do produce a better fit and more accurate approximations of the viscosity of ethane over the given temperature range (the boiling point of ethane is $-88.5^\circ C$).</p>

## Sanity Checking the Model 
<p>Now that we have reproduced the published coefficients, we can compare our results against the experimental results published by the authors.</p>
<p>First, we need to define some helper functions that will aid in calculations.</p>

```{r}
T0_calc <- function(coeffs, N) {
  T0_approx = coeffs[1] + coeffs[2]*N + coeffs[3]*N^2 + coeffs[4]*N^3
  return (T0_approx)
}

B_calc <- function(coeffs, N) {
  B_approx = coeffs[1] + coeffs[2]*N + coeffs[3]*N^2  + coeffs[4]*N^3
  return (B_approx)
}

viscosity_calc <- function(T0, B, Temperature) {
  log_visc = B*(1/Temperature - 1/T0)
  # The log in the de Guzman eqn is base 10.
  return(10^(log_visc))
}

```

#### Viscosity Function: De Guzman - Andrade
<p>Once the regression coefficients have been determined, we can use them to calculate values of viscosity v. temperature. The equation that is used for this is Eq 6 (page 7), the De Guzman-Andrade equation.</p>
<p>$log(\eta_L) = B\biggl[\frac{1}{T} - \frac{1}{T_0} \biggr]$</p>
<p>Note that the log in the above equation is base 10, not base e and the temperatures must be in the Kelvin scale NOT Celsius.</p>

```{r}
T_coeffs = coef(fit_cubic_T0)
B_coeffs = coef(fit_cubic_B)

alkane_names = c("Methane","Ethane", "Propane", "Butane", "Pentane", "Hexane", "Heptane", "Octane", "Nonane", "Decane",
                 "Undecane", "Dodecane", "Tridecane", "Tetradecane", "Pentadecane")

for (i in 1:15) {
  filename = paste0("data/",i,"-temp-viscosity.csv")
  alkane_tbl = read_csv(filename, col_names = TRUE)
  
  # generate the constants for the current alkane 
  T0_alkane = T0_calc(T_coeffs, i)
  B_alkane = B_calc(B_coeffs,i)
  
  # Add a new column with temps in Kelvin for use in calculations
  alkane_tbl = alkane_tbl %>% 
    mutate(TempK = alkane_tbl$Temp + KELVIN)
  
  # Compute viscosity
  alkane_tbl = alkane_tbl %>% 
    mutate(CalcTest = viscosity_calc(T0_alkane, B_alkane, alkane_tbl$TempK))
  
  plot_title = paste0("Viscosity of ", alkane_names[i], " v. Temperture (C)") 
  
  print(alkane_tbl %>% 
    ggplot(aes(x=Temp)) +
    geom_point(aes(y=Exp, color="Experimental")) +
    geom_line(aes(y=Exp, color="Experimental")) +
    geom_point(aes(y=Calc, color="Calc Published")) +
    geom_line(aes(y=Calc, color="Calc Published")) +
    geom_point(aes(y=CalcTest, color="Calculated")) +
    geom_line(aes(y=CalcTest, color="Calculated")) +
    theme_bw() +
    theme(legend.position="right") +
    labs(x="Temperature (C)", y="Viscosity", title=plot_title) +
    scale_color_manual(name="Viscosity Curves", values = c("Experimental"="#000000","Calculated"="#993399", "Calc Published"="#ff4500")))
  
}
```

```{r}
filename = paste0("data/",2,"-temp-viscosity.csv")
ethane_tbl = read_csv(filename, col_names = TRUE)


# Add a new column with temps in Kelvin for use in calculations
ethane_tbl = ethane_tbl %>% 
  mutate(TempK = ethane_tbl$Temp + KELVIN)


fit = lm(log10(ethane_tbl$Exp) ~ ethane_tbl$TempK)
summary(fit)
plot(log10(ethane_tbl$Exp) ~ ethane_tbl$TempK, ylab='log of viscosity', xlab='Temp (K)')
abline(fit)
```
### TODO:
1.) inverse of temp diff for methane
2.) higher order fits (higher order for bigger molecules)
3.) plot fits for N > 10 alkanes
